%
%RESULTS Rev1
%RUNTIME
%SUCCESS/ERROR

%%Convergence of results? Not with memory, with configs though...
%After contacting the vendors with our initial results one of the parties suggested to demonstrate the optimal operation of their database. 
%This was formalized by sending out a Request-For-Information (RFI), specifying the benchmarks we were planning to run.
%3 out of 4 vendors chose to participate in the RFI, which resulted in an \emph{Optimized} configuration. 

In Fig.~\ref{fig:Fig02_WatdivVerticalScaling}, the rightmost panel corresponds to running the benchmark with the \emph{RFI-optimized} configuration.

%
%median/mean
%Bla_N1_64_W1000_Def     2.825023 	6.741764
%Gra_N1_64_W1000_Def    40.375218 	59.564284
%Vir_N1_64_W1000_Def     0.175917 	3.667321
%
%Bla_N1_64_W1000_Opt    0.739366 	1.945429
%Gra_N1_64_W1000_Opt    0.652474  	6.320693
%Vir_N1_64_W1000_Opt    0.167420 	4.054050
%
\begin{itemize}
	\item \textbf{Sensitivity to configuration:} \textbf{Vir1\_64} did not benefit from the RFI settings file. For \textbf{Bla1\_64} the only improvement was to explicitly configure the timeout parameter on the server side. This avoids unnecessary overhead while the client is already disconnected. It leads to a speedup of approximately 3.5 for both runtime measurements. \textbf{Gra1\_64} benefits most from proper configurations: The provided scripts yielded a speedup of 9.4 for the average runtime and 62 for the median runtime.
	\item \textbf{32\_Doc to 64\_RFI:} From Fig.~\ref{fig:Fig02_WatdivVerticalScaling} it follows that, given sufficient memory and proper configuration, all stores have similar performance.
		%Moving from the left panel to the right in , we clearly see results converging in the rightmost window with the \textbf{64\_Opt} measurements. 
	\item \textbf{Average vs.\ median runtime:} \textbf{Bla1\_64} has an average runtime of 1.95s per query, whereas the measurements of \textbf{Vir1\_64} and \textbf{Gra1\_64} are 4.05s and 6.32s for  respectively. The median runtimes for \textbf{Vir1\_64} are the lowest at 0.17s where \textbf{Gra1\_64} and \textbf{Bla1\_64} have runtimes of 0.65s and 0.74s respectively.
%Bla_N1_32_W100_Def   0.191359		0.461707	=> x5
%Gra_N1_32_W100_Def   0.047812		0.417214 	=> x15
%Vir_N1_32_W100_Def   0.046707		0.301350	=> x15
%	\item \textbf{Runtime vs.\ Dataset Size:} Returning to section~\ref{subsec:bigdata} we can verify that the linear scaling behavior is largely restored, confirming our earlier hypothesis. Multiplication factors drop to 4.2 for Blazegraph, for Virtuoso and GraphDB mf $ \approx 15$.
\item \textbf{Timeouts \& Errors:} Apart from 5\% timeouts for \textbf{Gra1\_64\_RFI}, no query errors are observed with the \emph{RFI-Optimized} configurations.
\end{itemize}
%Bla_N1_64_W1000_Opt:	Success: 100.	Error: 0.0	Timeout: 0.0
%Gra_N1_64_W1000_Opt:	Success: 95.0	Error: 0.0	Timeout: 5.0
%Vir_N1_64_W1000_Opt:	Success: 100.	Error: 0.0	Timeout: 0.0

